discussion

### 2018-01-24 analysis/rwi/climate_rwi_relate.Rmd
update

>>>>

# SELECTED MONTHS

* Tras comentar con GGea, decidimos, poner desde sep-1 hasta oct1 
* quitar tmean

```{r}
set.seed(3333)
rprec_SJ <- dcc(chrono = cro_sj, climate = prec_N, 
         method = "correlation",  boot = "std",
        .range(-9:10) + .sum(-9:8))

rprec_caL <- dcc(chrono = cro_caL, climate = prec_S, 
         method = "correlation",  boot = "std",
        .range(-9:10) + .sum(-9:8))

rprec_caH <- dcc(chrono = cro_caH, climate = prec_S, 
         method = "correlation",  boot = "std",
        .range(-9:10) + .sum(-9:8))

## Plot precipitation 
```{r rwi_clim_prec_selec}
p_prec <- rprec_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rprec_all$id, labels = rprec_all$month_name) + 
  theme(legend.position = "none") + 
  annotate("text", label = "Prec", x = 1, y= -0.4) 
p_prec

pdf(paste0(di, '/out/climate_rwi/prec_selec.pdf'), width=9, height = 5)
p_prec
dev.off()
```

```{r}
set.seed(3333)
rsequia_SJ <- dcc(chrono = cro_sj, climate = sequia_sj, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rsequia_caL <- dcc(chrono = cro_caL, climate = sequia_caL, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rsequia_caH <- dcc(chrono = cro_caH, climate = sequia_caH, 
         method = "correlation",  boot = "std", var_names = 'spei6',
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

## Plot SPEI HR
```{r rwi_clim_sequia_selec}
p_sequia <- rsequia_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rsequia_all$id, labels = rsequia_all$month_name) + 
  theme(legend.position = "none") + 
  annotate("text", label = "SPEI 6 - HR", x = 1.5, y= -0.4) 
p_sequia

pdf(paste0(di, '/out/climate_rwi/sequia_selec.pdf'), width=9, height = 5)
p_sequia
dev.off()
```

```{r}
set.seed(3333)
rtmin_SJ <- dcc(chrono = cro_sj, climate = tmin_N, 
         method = "correlation",  boot = "std", 
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmin_caH <- dcc(chrono = cro_caH, climate = tmin_S, 
         method = "correlation",  boot = "std", 
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmin_caL <- dcc(chrono = cro_caL, climate = tmin_S, 
         method = "correlation",  boot = "std", 
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

## Plot Tmin 
```{r rwi_cli_tmin_selec}
p_tmin <- rtmin_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rtmin_all$id, labels = rtmin_all$month_name) + 
  theme(legend.position = c(.1, .8)) +
  annotate("text", label = "Tmin", x = 1, y= -0.4) + ylim(-.4, .4)
p_tmin

pdf(paste0(di, '/out/climate_rwi/tmin_selec.pdf'), width=9, height = 5)
p_tmin
dev.off()
```

```{r}
set.seed(3333)
rtmax_SJ <- dcc(chrono = cro_sj, climate = tmax_N, 
         method = "correlation",  boot = "std", 
        .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmax_caH <- dcc(chrono = cro_caH, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

rtmax_caL <- dcc(chrono = cro_caL, climate = tmax_S, 
         method = "correlation",  boot = "std", 
         .range(-9:10) + .mean(-9:8) + .mean(-12:2) + .mean(3:5) + .mean(6:8) + .mean(9:11))

## Plot Tmax 
```{r rwi_cli_tmax_selec}
p_tmax <- rtmax_all %>% ggplot(aes(x=id, y=coef, group = site)) + 
  custom_gg +
  scale_x_continuous(breaks = rtmin_all$id, labels = rtmin_all$month_name) +
  theme(legend.position = "none") +
  annotate("text", label = "Tmax", x = 1, y= -0.4) + ylim(-.4, .4)
p_tmax

pdf(paste0(di, '/out/climate_rwi/tmax_selec.pdf'), width=9, height = 5)
p_tmax
dev.off()
```

```{r rwi_cli_relation}
p <- grid.arrange(p_tmin, p_tmax, p_prec, p_sequia, ncol=2)
pdf(paste0(di, '/out/climate_rwi/rwi_cli_relation.pdf'), width=9, height = 12)
p
dev.off()
```

### 2018-01-24 man/ms_1_material_methods.Rmd
update

### Tree sampling 
Sampling was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern aspect) and Cáñar (CA; southern aspect) (Figure 1; Table 1). For the southern site two elevations were sampled: CA-Low and CA-High. All the sites were oak monospecific and representatives of two of the three population clusters identified for the species in this mountain range [@PerezLuque2015]. In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each tree at breast heigth (1.3 m) using an increment borer. Diameter at breast height (DBH) and total height were measured using a girth tape and a Vertex IV ultrasonic hypsometer (Haglöf, Sweden) respectively. We assessed stand competition of target trees by recording distance, azimuth, dbh, species and total height of all neighboring living trees with dbh > 7.5 cm in a circular plot of 10 m radius. We used non-parametric Kruskal-Wallis rank sum tests [@Sokal1995] for search differences between sites for height, DBH and competence indices. When significant differences were observed, we run multiple comparisons using the Dunn's-test [@Dunn1964] with Bonferroni adjustment to correct significance values. These analysis were performed uisng the `PMCMR` package [@Pohlert2014].

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and mean, minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. Data were downloaded and preprocessed using the climate4R bundle (http://www.meteo.unican.es/en/climate4R) [@Cofino2018; @Frias2018].
We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010; @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We selected a temporal scale for the SPEI of 6 months.

>>>>

### Tree sampling 
Sampling was carried during autumn of 2016. Trees were sampled at two locations in contrasting slopes of Sierra Nevada: San Juan (SJ; northern aspect) and Cáñar (CA; southern aspect) (Figure 1; Table 1). For the southern site two elevations were sampled: CA-Low and CA-High. All the sites were oak monospecific and representatives of two of the three population clusters identified for the species in this mountain range [@PerezLuque2015]. In each site between 15 and 20 dominant trees were randomly selected. Two cores of 5 mm of diameter were taken from each tree at breast heigth (1.3 m) using an increment borer. Diameter at breast height (DBH) and total height were measured using a girth tape and a Vertex IV ultrasonic hypsometer (Haglöf, Sweden) respectively. We assessed stand competition of target trees by recording distance, azimuth, dbh, species and total height of all neighboring living trees with dbh > 7.5 cm in a circular plot of 10 m radius. We computed two distance independent indices: density in $\mathrm{trees \cdot ha^{-1}}$, and basal area (BA) in $\mathrm{m^{2} \cdot ha^{-1}}$; and one distance dependent indices: size ratio proportional to distance (*srd*) as $\mathrm{srd} = \sum_{i=1}^{n} (dbh_j / dbh_i) \cdot \left[1/ (dist_{ij} + 1) \right ]$. We used non-parametric Kruskal-Wallis rank sum tests [@Sokal1995] for search differences between sites for height, DBH and competence indices. When significant differences were observed, we run multiple comparisons using the Dunn's-test [@Dunn1964] with Bonferroni adjustment to correct significance values. These analysis were performed uisng the `PMCMR` package [@Pohlert2014].

@@

Climate data were obtained from the European Daily High-Resolution Observational Gridded Dataset (E-OBS v16) [@Haylock2008]. Monthly precipitation and minimum and maximum temperatures were obtained at a 0.25 x 0.25 º resolution for the 1950-2016 period. We selected grid cells covering each sampled sites. Data were downloaded and preprocessed using the climate4R bundle (http://www.meteo.unican.es/en/climate4R) [@Cofino2018; @Frias2018]. We also used the Standardized Precipitation-Evapotranspiration Index (SPEI), a multiscalar drought index that incorporates both precipitation and temperature [@VicenteSerrano2010; @Begueria2014]. SPEI values for the period 1961-2014 were obtained with a spatial resolution of 1.1 km from the Drought indices dataset for Spain database (http://monitordesequia.csic.es/), a high resolution database of drought indices for Spain [@VicenteSerrano2017]. We selected a temporal scale for the SPEI of 6 months.

### 2018-01-24 man/ms_2_results.Rmd
update

The trees of the southern population were older and larger than northern ones, particularly for the highest elevation site (Ca-High) which had bigger trees than the other sites (Tables 1, 2). The basal area increment (BAI) were different between sites (Figure 5). El sitio mas elevado (Ca-High) mostró mayores valores de BAI que los otros sitios, mostrando además una tendencia positiva a partir de 1970.

>>>>

The trees of the southern population were older and larger than northern ones, particularly for the high elevation site (Ca-High) which had bigger and taller trees than the other sites (Tables 1, 2). Competence indices were similar among sites (Table 1). The basal area increment (BAI) were different between sites (Figure 5). El sitio mas elevado (Ca-High) mostró mayores valores de BAI que los otros sitios, mostrando además una tendencia positiva a partir de 1970. Se observan diferencias en cuanto al crecimiento entre los sitios de las poblaciones del sur. Ca_Low se parece en crecimiento mas a SJ que a Ca_High.

### 2018-01-24 man/ms_3_discussion.Rmd
update

>>>>

## Dendro

### Diferencias crecimiento entre sitios. 
Mayor crecimiento en los robles situados en CaHigh. Sorprende que en el rear edge de la distribución de la especie, en la cara sur dentro de este rear edge, y en la parte mas alta (estas parcelas están en el treeline de la especie en SN) es donde encontramos mayores crecimientos. Parece que los robles estuvieran mejor en esta zona. Algunas reflexiones

* Quizá están creciendo donde les están dejando (menor impacto antrópico)
* ¿Menor estrés hídrico en zonas mas altas? 
* ¿Diferencias entre suelos? (ver artículo [@CoboDiaz2017] que compara suelos de Cáñar). En este estudio sobre microbiota del suelo, analizó tres sitios: por encima del treline (XZF), low altitudinal oak forest (LAF) and high altitudinal oak forest (HAF). LAF está mucho mas bajo, pero HAF está cerca de las parcelas caHigh y encontró lo siguiente:

  * tipo de suelo: LAF es Sandy-loam, HAF es loam
  * HAF tuvo los valores mayores de disponibilidad hídrica (% available water tabla s1 en [@CoboDiaz2017]) 
  * Los tres sitios eran pobres en materia orgánica, pero el HAF dobló los valores de los otros dos sitios
  * No diferencias en contenido de N, C/N ratio similares en HAF, y LAF

Sorprende que en Cáñar haya diferencias tan grandes de crecimiento estándo tan cerca. ¿Puede estar relacionado con intensidad de uso antrópico? De hecho la Ca-Low tiene crecimientos similares a SJ.
